RNeXML: a package for reading and writing richly annotated phylogenetic, character, and trait data in R

### 2014-11-28 manuscripts/manuscript.Rmd
Smoothes and slightly tightens Intro to RNeXML leader.

A NeXML document can be thought of as having the following components: (1) phylogeny
data, (2) character-trait data, (3) operational taxonomic units, and (4) metadata. 
`RNeXML` seeks to provide a seamless way to convert these to and from their 
corresponding R object representations, in particular `ape::phylo`,
`data.frame`, and lists for phylogenies, character
matrices, and metadata, respectively.
Metadata can provide bibliographic and provenance
information about the study as part of which the phylogeny was generated or
applied (such as authors, citations to publications, timestamps, and even
descriptions of methods). They can also be used to attach annotations to individual
data elements.  We illustrate how metadata annotations can be added to
the operational taxonomic units of the phylogeny and can be used to extend
the data representation to new uses such as stochastic character maps.
<!-- sounds awkward - need to smoothen the language. HL -->

>>>>

Conceptually, a NeXML document has the following components: (1) phylogeny
data, (2) character or trait data in matrix form, (3) operational
taxonomic units (OTUs), and (4) metadata. When reading and writing NeXML,
`RNeXML` aims to map these to and from, respectively, their most
natural representations in R. These are the `phylo` object from the
`ape` package [@Paradis_2004], R's native `data.frame` list structure for
phylogenies, character matrices, and metadata, respectively.
<!-- compared to the preceding enumeration, this is missing where the OTUs go -->

NeXML allows attaching metadata to any data element, and even to
metadata themselves. At the level of the document as a whole or the
phylogeny, metadata can provide bibliographic and provenance
information, for example about the study as part of which the
phylogeny was generated or applied, and with which methods it was
generated. In the form of annotations, metadata can also be attached
to very specific elements of the data, such as individual OTUs, nodes,
or even edges of the phylogeny.
